---
title: "Final Project - GEOG 533"
author: "Daniella Madubuike"
date: "December 3, 2017"
output: html_document
---

###Changing Car Culture Through Carpooling
#####Description
Research focused on measuring the carpooling potential of students by estimating the overall interest and willingness of students to utilize this potential alternative transportation option.

#####Need for Analysis
To compare variables obtained from survey and find correlations and positive/negative relationships.

#####Problem
Binghamton Univeristy has a large parking problem. It also lacks student location data.

#####Results/Benefits
The tool can be incorporated into potential future carpooling systems.  Could help to reduce the inherent car culture of single-occupancy driving. 

#####Test Data 
Survey questions based on satisfaction, drive time, comfort level with carpooling, perceived benefits, and perceived drawbacks of carpooling, were asked, in both open-ended and close-ended question format.



##Location Analysis
####Distance Based Analysis: Average Nearest Neighbor 

```{r}
#####Change working directory to where data below is located.
setwd("/Users/daniellamadubuike/Documents")
library(ggmap)
library(Rcpp)

if(!require(readxl)) install.packages("readxl")
library(readxl)
lats <- read_excel("excellatlonwo.xlsx")

data(lats)
plot(lats$Longitude,lats$Latitude,xlab="Longtitude",ylab="Latitude")
```


Using the 'spatstat' package.
```{r, message=FALSE}
if(!require(spatstat)) install.packages("spatstat")
library(spatstat)
P <- lats
```

To compute the average *first* nearest neighbor distance (1st order):
```{r}
mean(nndist(P, k=1))
```

To compute the average *second* nearest neighbor distance (2nd order):
```{r}
mean(nndist(P, k=2))
```

To plot average distance as a function of neighbor order for the first 100 closest neighbors:
```{r}
ANN <- apply(nndist(P, k=1:100),2,FUN=mean)
plot(ANN ~ eval(1:100), type="b", main=NULL )
```

The bottom axis shows the neighbor order number and the left axis shows the average distance.



##Statistical Analysis
####Categorical Data: Chi – Square Test

Public Transportation Frequency *VS* Satisfaction With Parking

Null: There is no correlation between student public transportation use frequency and their satisfaction with parking
Alternative: There is a correlation between student public transportation use frequency and their satisfaction with parking

```{r,message=FALSE}
if(!require(readxl)) install.packages("readxl")
library(readxl)
excel <- read_excel("excel.xlsx")
tbl = table(excel$`Public transportation Use Frequency`, excel$`Satisfaction W Parking`)
tbl
chisq.test(tbl, simulate.p.value = TRUE)
```

P-value is 0.04, we reject the null.

#####Continuous Data: Anova Test
Income *VS* Parking Time

Null: There is no correlation between student income and the amount of time is takes them to park.
Alternative: There is a correlation between student income and the amount of time is takes them to park.
```{r}
if(!require(readxl)) install.packages("readxl")
library(readxl)
excel <- read_excel("excel.xlsx")

x <- excel$Income
y <- excel$`Parking Time`
plot(y ~ x)
m <- lm(y~x)
summary(m)
abline(m,col="red",lwd=2)
```

P-value is 0.539, so we fail to reject the null.

##Qualitative Data
####World Cloud
Data Mining
```{r, message=FALSE}
if(!require(tm)) install.packages("tm")  # for text mining
if(!require(SnowballC))install.packages("SnowballC") # for text stemming
if(!require(wordcloud))install.packages("wordcloud") # word-cloud generator 
if(!require(RColorBrewer))install.packages("RColorBrewer") # color palettes

library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
```

Place data in text file and read into R, then create Corpus
```{r}
filepath <- "https://dmadubu1.github.io/comment.txt"
text <- readLines(filepath)
docs <- Corpus(VectorSource(text))
#inspect(docs)
```

Remove unwanted punctuation marks
```{r}
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")
```

Remove other unwanted words and others
```{r}
# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove your own stop word
# specify your stopwords as a character vector
docs <- tm_map(docs, removeWords, c("blabla1", "blabla2")) 
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
# docs <- tm_map(docs, stemDocument)
```

Sort text file
```{r}
dtm <- TermDocumentMatrix(docs)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)
```
 
Set seed and creat world cloud
```{r}
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=150, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(6, "Dark2"))
```

World cloud show trend words like parking, campus, spots, students, people, carpooling, incentive. As expected.



