---
title: "Geog533 Lab 8"
author: "Daniella Madubuike"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|     0.6         |        1            |    0.6        |        7.79     |
| Residual |       1.7       |       22            |   0.077       |                 |
| Total    |       2.3       |        21           |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
#y = mx + b
y = 0.19*50000 + 0.46

#to convert data
ans <- y/1000
ans
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

####The coefficient 0.19 is the slope, which is the measure of the steepness of the line. It indicates a change in y after a unit change in x.


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
qf(0.95, 1, 22)
```
####The regression coefficient is Not significantly diff from zero because the f stat (4.3) is less than the f crit (7.79) so we accept the null of equal to zero.


### Question 1(e)
What is the value of the correlation coefficient?
```{r}
#r <- ssm / sst
r <- (2.3 - 1.7)/ 2.3
r
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
s <- c(36, 78, 11, 45)
e <- c(400, 800, 200, 675)

 m <- lm(s ~ e)
 summary(m)
 anova(m)
```
####Intercept coefficient is -7.63721, slope coefficient is 0.0966

### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
```
####11.25 on 2 degrees of freedom

### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
qf(0.95, 1, 2)
```
####Since F -statistical(16.17) is less than F critical (18.51), we accept the null.

### Question 2(d)
Find the value of $r^2$.
```{r}
```
####R squared = 0.8899

### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
qo <- data.frame(anova(m))
qo
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
aov(m)
anova(m)
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|      2047.77    |        1           |     2047.77    |      16.17      |
| Residual |       253.22    |        2           |     126.61     |                 |
| Total    |      2300.99    |        3           |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(e, s)
abline (m, col="red")
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
url <- "Milwaukee_Sales.csv"
df <- read.csv(url,header = TRUE)
m <- lm(df$SalePrice ~ df$LotSize)
summary(m)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
url <- "UK_Housing.csv"
df <- read.csv(url,header = TRUE)
m <- lm(df$price ~ df$bedrooms)
summary(m)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
url <- "UK_Housing.csv"
df <- read.csv(url,header = TRUE)
m <- lm(df$price ~ df$bathrooms)
summary(m)
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
if(!require(leaflet)) install.packages("leaflet")
url2 = "http://spatial.binghamton.edu/geog533/data/hometown.csv"
url2 <- "hometown.csv"
df <- read.csv(url2, header = TRUE)
m <- leaflet()
m <- addTiles(m) # Add default OpenStreetMap map tiles
m <- addMarkers(m, lng=df$latitude, lat=df$longitude, popup=df$fullname)
m
```

