---
title: "Geog533 Lab 10"
author: "Daniella Madubuike"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
### install packages
if(!require(UScensus2010)) install.packages("UScensus2010")
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx")
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010county)
library(UScensus2010tract)
data(new_york.county10)
shp <- new_york.county10
plot(shp)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
if(!require(tmap)) install.packages("tmap")
library(tmap)
qtm(shp)
qtm(shp,fill = "NAME10")
qtm(shp,fill = "P0010001")
```


### Question 1(c)	
How many counties in New York State?
```{r}
count <- shp$county
length(table(count))
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
#if (shp$NAMELSAD10 == "Broome County") {
  #print(shp$county)}

boom <- shp$county == "Broome County"
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
if(!require(fBasics)) install.packages("fBasics")
library(fBasics)
pop <- shp$P0010001

basicStats(pop)
summary(pop)

```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
if(!require(UScensus2010)) install.packages("UScensus2010")
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx")
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010county)
library(UScensus2010tract)
data(new_york.tract10)
shp <- new_york.tract10
df <- shp@data
plot(shp)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df <- shp@data
sum(pop)
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
library(rgdal)
df <- shp@data
sel <- data.frame(df$county == "007")
shp[sel, ]
plot(shp[sel,])
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
sum(sel)
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
df <- shp@data
sel <- df$county == "007"
shp[sel, ]
hist(sel)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}

```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
#setwd("Downloads/")
#getwd()
#if(!require(ncdf4)) install.packages("ncdf4")
#library(ncdf4)
#library(raster)
#ndvi.rb <- brick("GIMMS.NDVI.360.720.2000.2002.30days.nc")
#ndvi.rb

if(!require(ncdf4)) install.packages("ncdf4")
library(ncdf4)
library(rgdal)
library(raster)
multi <- raster("/Users/daniellamadubuike/Downloads/NDVI.nc/")
multi
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(multi)
ncol(multi)
ncell(multi)
extent(multi)
bbox(multi)
res(multi)
projection(multi)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
meanmulti <- as.data.frame(cellStats(multi2,mean))
meanmulti
maxmulti <- as.data.frame(cellStats(multi2,max))
maxmulti
writeRaster(meanmulti, filename = "meanmulti.grd")
writeRaster(maxmulti, filename = "maxmulti")
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
multi2 <- brick("/Users/daniellamadubuike/Downloads/NDVI.nc/")
multi2
plot(multi2, 13:24)


```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r, warning=FALSE}
hist(multi2, 13:24)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 

```{r}
plot(multi2, 7)
values <- click(multi2, n=1, xy=TRUE)
#values <- click(multi2, n=1, xy=FALSE)
plot(as.vector(values), type="b")
```

