---
title: "Lab 4"
author: "Daniella Madubuike"
date: "October 9, 2017"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Question 1a
The number of points scored by each team in a tournament is normally distributed, with
mean μ = 32 and standard deviation σ = 7. What is the probability of: A team scoring 20 points or fewer? Plot the density function and shade the area. 
```{r}
curve(dnorm(x, mean = 32, 7), from = 0, to = 53, main = '20 Points or Fewer')
cord.x <- c(0, seq(0,20,0.01), 20)
cord.y <- c(0, dnorm(seq(0,20,0.01), mean = 32, sd = 7), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 1b
A team scoring more than 35 points? Plot the density function and shade the area.
```{r}
curve(dnorm(x, mean = 32, 7), from = 11, to = 53, main = 'More than 35 Points')
cord.x <- c(35, seq(35,53, 0.01), 53)
cord.y <- c(0, dnorm(seq(35,53,0.01), mean = 32, sd = 7), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 1c
A team scoring between 20 and 40 points? Plot the density function and shade the area. 
```{r}
curve(dnorm(x, mean = 32, 7), from = 11, to = 53, main = 'Between 20 and 40 points')
cord.x <- c(20, seq(20,40, 0.01), 40)
cord.y <- c(0, dnorm(seq(20,40,0.01), mean = 32, sd = 7), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 2a
The number of comments per post on a social media site is exponentially distributed, with the
average post receiving ten comments. What percentage of posts get: Fewer than three comments? Plot the density function and shade the area.

```{r}
x <- dexp(3, rate = 0.1)
curve(dexp(x, 0.1), from = 0, to = 90, main = 'Fewer than 3 Comments')

cord.x <- c(0, seq(0,3,0.01), 3)
cord.y <- c(0, dexp(seq(0,3,0.01), 0.1), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 2b
More than 20 comments? Plot the density function and shade the area
```{r}
x <- pexp(20, rate = 0.1, lower.tail = FALSE) 
curve(dexp(x, 0.1), from = 0, to = 90, main = 'More than 20 Comments')

cord.x <- c(20, seq(20,90,0.01), 90)
cord.y <- c(0, dexp(seq(20,90,0.01), 0.1), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 2c
Between five and ten comments? Plot the density function and shade the area. 
```{r}
curve(dexp(x,0.1), from = 0, to = 90, main = 'Between 5 and 10 Comments')

cord.x <- c(5, seq(5,10, 0.01), 10)
cord.y <- c(0, dexp(seq(5,10,0.01), 0.1), 0)
polygon(cord.x, cord.y, col = "red")
```

###Question 3a
Basic raster creation and calculation: Create a first raster (30 rows × 30 columns) and assign values to the raster based on
random values from a uniform distribution. Plot the raster and the histogram.
```{r, message=FALSE, warning=FALSE}
if(!require(raster)) install.packages("raster")
library(raster)
ras <- raster(nrows=30,ncols=30,xmn=0,xmx=10,ymn=0,ymx=10)
set.seed = 100
ras[] <- runif(ncell(ras))
plot(ras)
hist(ras)
```

###Question 3b
Create a second raster (30 rows × 30 columns) and assign values to the raster based
on random values from a normal distribution. Plot the raster and the histogram.
```{r}
ras2 <- raster(nrows=30,ncols=30,xmn=0,xmx=10,ymn=0,ymx=10)
set.seed = 200
ras2[] <- rnorm(ncell(ras2))
plot(ras2)
hist(ras2)
```

###Question 3c
Create a third raster (30 rows × 30 columns) and assign values to the raster based on
cell-by-cell addition of the first raster and the second raster. Plot the raster and the
histogram.

```{r}
newraster <- ras + ras2
plot(newraster)
hist(newraster)
```

###Question 3d
Calculate the mean value of the third raster. Reclassify the third raster into a binary
image: 1 (cell value > mean value) and 0 (cell value <= mean value). Save the
reclassified image to your working directory as a TIFF image.
```{r}
a <- cellStats(newraster,"mean")
newraster[newraster > a ] <- 1
newraster[newraster <= a] <- 0
plot(newraster)
writeRaster(newraster,filename = "newraster.tif", overwrite = TRUE)
read <- raster("newraster.tif")
```